An Extensible Approach to Verification of Embedded Network Systems*

### 2014-04-25 tech-project/implementation.tex
tweaked and added info to the start of the memory allocation section

% IN THIS SECTION SHOW SOME OF OUR TRANSLATED CODE... AND WALK THROUGH IT THE SAME WAY YOU DID WITH THE CONCEPT IN RESOLVE. BUT BEFORE YOU DO THAT, SHOW THE PICTURE (The one thats already here showing high level relationships and update it!).
\section{Implementation}
Development of our C translation tool can be logically partitioned into three distinct phases: 
\begin{enumerate}
\item Arriving at a translation model (or, strategy) for an accurate C representation of RESOLVE.
\item Implementing reusable mechanisms for carrying out the C code generation process.
\item Creation of a memory manager capable of safely allocating and freeing dynamic memory required by the generated code.
\end{enumerate}
We illustrate each of these phases working in tandem on the LED component discussed in Section \ref{sec:specifiying}.

\subsection{C Translation Model}
One of the primary challenges in translating from RESOLVE to C is finding a suitable C analog for each RESOLVE module and the constructs allowable in each. Indeed, since we are dealing with an environment where functional correctness is a primary concern, it is important that the code generated by our tool represents as closely as possible the original RESOLVE source. In an effort to make such considerations, at the highest level, the C code we generate makes special considerations for facilities, concepts, and realizations. This scheme is depicted in Figure \ref{fig:relationship} and briefly summarized in Sections \ref{sec:conceptoverview} - \ref{sec:facilitiesrealizations}.

@@

\subsubsection{Concepts}
\label{sec:conceptoverview}
Concept modules produce a single .h file, which provides function pointers for the operations specified in the original concept, as well as structs representing each user defined type.

\subsubsection{Facilities}
\label{sec:facilitiesoverview}
Facilities produce an .h/.c pair: The header .h declares methods that are intended to manage the creation and destruction of all global variables used within the facility, while the .c provides an implementation of these methods.  Note that these create and destroy methods are only responsible for freeing \textit{global} variables. All other translated local functions are responsible for deallocating their own local variables.

@@

\begin{figure*}
\begin{center}
\includegraphics[scale=.55]{figs/ast_traversal2.pdf}
\end{center}
\caption{The general flow of information from C translators walk of the AST (first), to user defined templates (middle), ending with formed output (last).}
\label{fig:ast}
\end{figure*}

@@

The C code generation process takes place over the course of a traversal of RESOLVE's abstract syntax tree (AST). The traversal mechanism used is a derivative of the visitor pattern that provides a pre post traversal over all nodes in the tree.

@@

%%LOOK AT ME
%This separation between the RESOLVE source tree and the view, allows us to easily modify and tweak the code produced by our tool and allows us easily represent arbitrarily complicated output

Shown in Figure \ref{fig:ast} is a high level depiction of steps taken in translating this operation to C. The first box depicts the AST of \texttt{Nothing}, where nodes are represented with boxes labeled by the constructs they contain. Throughout the walk of this tree, useful information (such as the operation's name, ``Nothing") are extracted from the nodes containing these constructs, and added to the externally defined string-template\footnote{A template can simply be thought of as a ``document with holes" which the user choses when and how to fill}, \texttt{c\_function\_def} template.

@@

We feel this approach lends itself the challenge discussed in this paper, as this separation allows us to easily iterate changes to our generated C code without needing to concern ourselves with a line of Java inside the compiler itself. This allows us to easily tweak code and make arbitrarily complicated changes and optimizations to the code -- some of which are detailed later in Section \#.

@@

Dynamic memory allocation is not typically the first choice for most embedded applications. With the telos mote itself  limited to 128 bytes of RAM, hardware and memory constraints such as these, developers of embedded applications tend to favor static memory allocation over dynamic. Many programs, however benefit greatly from dynamic memory allocation not only in terms of clarity and straightforward to be used, including RESOLVE, in order to make it extensible. Previous RESOVE translations to C required static memory allocation \cite{regula:2010}. In this section a stack based dynamic memory allocator is introduced.

@@

The \texttt{salloc()} is a first fit memory allocator. Rather than allocating on the heap, \texttt{salloc()} uses the stack. At compilation, the allocator provisions a fixed size of memory. It requires a small section of meta-data called a block which contains information of about the size of memory allocated, neighboring blocks, as well if the block is free or not. A sample representation of the stack shown in figure \ref{fig:stack}, is a typical example of allocated memory on the stack.

>>>>

% IN THIS SECTION SHOW SOME OF OUR TRANSLATED CODE... AND WALK THROUGH IT THE SAME WAY YOU DID WITH THE CONCEPT IN RESOLVE. BUT BEFORE YOU DO THAT, SHOW THE PICTURE (The one thats already here showing high level relationships and update it!).
\section{Implementation}
Development of our C translation tool can be logically partitioned into three distinct phases: 
\begin{enumerate}
\item Arriving at a translation model (or, strategy) for an accurate C representation of RESOLVE.
\item Implementing reusable mechanisms for carrying out the C code generation process.
\item Creation of a memory manager capable of safely allocating and freeing dynamic memory required by the generated code.
\end{enumerate}
We conclude this section with a demonstration of each of these phases working in tandem on the LED component discussed in Section \ref{sec:specifiying}.

\subsection{C Translation Model}
One of the primary challenges in translating from RESOLVE to C is finding a suitable C analog for each RESOLVE module and the constructs allowable in each. Indeed, since we are dealing with an environment where functional correctness is a primary concern, it is important that the code generated by our tool represents as closely as possible the original RESOLVE source. In an effort to make such considerations, at the highest level, the C code we generate makes special considerations for concepts, facilities, and realizations. This scheme is depicted in Figure \ref{fig:relationship} and briefly summarized in Sections \ref{sec:conceptoverview} - \ref{sec:facilitiesrealizations}.

@@

\subsubsection{Concepts}
\label{sec:conceptoverview}
Concept modules produce a single .h file, which provide function pointers for the operations specified in the original concept, as well as structs representing each user defined type.

\subsubsection{Facilities}
\label{sec:facilitiesoverview}
Facilities produce an .h/.c pair: The header .h declares both a ``create" and ``destroy" method which, taken together, encapsulates the creation and destruction of all global variables used within the facility. The .c provides an implementation of these methods.  Note that these create and destroy methods are only responsible for freeing \textit{global} variables -- meaning all other translated functions are responsible for deallocating their own local variables.

@@

\begin{figure*}
\begin{center}
\includegraphics[scale=.55]{figs/ast_traversal2.pdf}
\end{center}
\caption{The general flow of information from the AST (first), to user defined templates (middle), ending with formed output (last).}
\label{fig:ast}
\end{figure*}

@@

Translation itself occurs over the course of a traversal of RESOLVE's abstract syntax tree (AST) -- an intermediate representation of RESOLVE code. The traversal mechanism utilized is a derivative of the visitor pattern that provides a pre post traversal over all nodes in the tree.

@@

Shown in Figure \ref{fig:ast} is a high level depiction of steps taken in translating this operation to C. The first box depicts the AST of \texttt{Nothing}, where nodes are represented with boxes labeled by the constructs they contain. Throughout the walk of this tree, useful information (such as the operation's name, ``Nothing") are extracted from the nodes containing these constructs, and added to a user defined string-template\footnote{A template can simply be thought of as a ``document with holes" which the user choses when and how to fill.}, in this case: \texttt{c\_function\_def}.

@@

We feel this approach lends itself well to the challenge discussed in this paper, as this separation allows us to easily iterate changes to our generated C code without needing to concern ourselves with the Java written inside the compiler itself. This allows us to easily and iteratively tweak generated code -- making arbitrarily complicated changes and optimizations, some of which are detailed later in Section \#.

@@

Any model seeking to convert RESOLVE to a lower level representation such as C demands a strategy for handling any memory utilized by the generated code. While dynamic allocation is typically the norm, it is not however the first choice for embedded applications. With the Telos mote constrained to 128 bytes of RAM, developers targeting embedded platforms tend to favor static memory allocation over dynamic, due the memory efficiency it affords.

%Many programs however benefit greatly from dynamic memory allocation not only in terms of clarity and straightforwardness, but in our case: Accurately modeling exactly what the code being executed is doing. 
%NesC, for example, while giving the illusion of dynamic allocation, behind the scenes actually performs static allocation

In spite of this, our reasoning for choosing to pursue dynamic allocation over static, rests in the language itself. Being a language that relies heavily on the notion of formally specified objects (components), everything in RESOLVE from the `primitive' types such as booleans and integers, to the more complicated (user defined) ones such as Stacks and Queues, are modeled, specified, and implemented in the same way: Using the standard RESOLVE machinery discussed in Section \ref{sec:specifiying}.

As such, even the simplest RESOLVE programs still rely heavily on the notion of objects coming and going. Thus, from a verification perspective, creating a tool that allows our translated code to create and destroy such objects in the manner the language itself uses is an important component to capture in formally verified code.

@@

The particular allocation scheme that we chose to use is \texttt{salloc()}: A first fit memory allocator. Rather than allocating memory on the heap, \texttt{salloc()} instead uses the stack. At compilation, the allocator provisions a fixed size of memory. It requires a small section of meta-data called a block which contains information of about the size of memory allocated, neighboring blocks, as well if the block is free or not. A sample representation of the stack shown in figure \ref{fig:stack}, is a typical example of allocated memory on the stack.

### 2014-04-25 tech-project/specify_led.tex
redid the translation implementation section
tweaked and added info to the start of the memory allocation section

\begin{figure*}[!htb]
\centering
\includegraphics[scale=.55]{figs/component_graph.pdf}
\caption{A diagram illustrating inter-component relationships for the example in Section \ref{sec:specifying}}
\end{figure*}
\label{fig:imp}

@@

\begin{figure*}
\centering
\includegraphics[scale=.55]{figs/ast_traversal.pdf}
\caption{A RESOLVE operation AST, the walk call sequence, and sample translation output.}
\end{figure*}
\label{fig:ast}

>>>>

\begin{figure*}[!htb]
\centering
\includegraphics[scale=.55]{figs/component_graph.pdf}
\caption{A diagram illustrating inter-component relationships for the example in Section 3. The files listed in cloud shapes indicate externally realized components, while the dotted line separating these from others indicates the current boundary between the lowest-level, hardcoded components in RESOLVE, and native RESOLVE abstractions sitting on top of these hardcoded components.}
\end{figure*}
\label{fig:imp}

@@
